---
author: "Suresh Gopalakrishnan"
code_folding: hide
date: "January 25, 2018"
linkcolor: cyan
number_sections: yes
output:
  html_document: default
  pdf_document: default
theme: cosmo
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Data Wrangling Exercise 2: Dealing with missing values
Suresh
Date- `r Sys.Date()` 

#### Load required packages 
```{r message=FALSE,warning=FALSE}
##some  packages may not be used. but loading basic DW pkgs.
library(dplyr) # DW pkg
library(tidyr) # DW pkg
library(data.table) # Base pkg for DescTools
library(DescTools) # For LIKE and ANY function
library(readr) # write faster csv files
```

#### 0: Load the data in RStudio - *"titanic_original.csv"*

```{r message=FALSE,warning=FALSE}
titanic <- read.csv("titanic_original.csv")
glimpse(titanic)
titanic <- data.frame(titanic)
```

#### 1: Port of embarkation

The embarked column has some missing values, which are known to correspond to passengers who actually embarked at Southampton. Find the missing values and replace them with S. (Caution: Sometimes a missing value might be read into R as a blank or empty string.)

```{r message=FALSE,warning=FALSE}
titanic <- titanic %>% mutate(embarked = ifelse(embarked == '','S',as.character(embarked)))
titanic$embarked <- as.factor(titanic$embarked)
```

#### 2: Age

You'll notice that a lot of the values in the Age column are missing. While there are many ways to fill these missing values, using the mean or median of the rest of the values is quite common in such cases.
Calculate the mean of the Age column and use that value to populate the missing values

```{r message=FALSE,warning=FALSE}
#Check number of NAs in age vector

sum(is.na(titanic$age))
    
#if NA update age with Mean of ages    

titanic <- titanic %>% mutate(age = ifelse(is.na(age), mean(age,na.rm=TRUE),age))

#Test all NA are removed

sum(is.na(titanic$age))
```

#### 3: Lifeboat

You're interested in looking at the distribution of passengers in different lifeboats, but as we know, many passengers did not make it to a boat :-( This means that there are a lot of missing values in the boat column. Fill these empty slots with a dummy value e.g. the string 'None' or 'NA'

```{r message=FALSE,warning=FALSE}
titanic <- titanic %>% mutate(boat = ifelse(boat == '','NA',boat))
```

#### 4: Cabin

You notice that many passengers don't have a cabin number associated with them.
Does it make sense to fill missing cabin numbers with a value?
What does a missing value here mean?
You have a hunch that the fact that the cabin number is missing might be a useful indicator of survival. Create a new column has_cabin_number which has 1 if there is a cabin number, and 0 otherwise.

```{r message=FALSE,warning=FALSE}
titanic$cabin <- as.character(titanic$cabin)
titanic <- titanic %>%
       mutate(has_cabin_number  = ifelse(cabin == '', 0, 1))
```

#### 5: Submit the project on Github

Include your code, the original data as a CSV file titanic_original.csv, and the cleaned up data as a CSV file called titanic_clean.csv.


```{r message=FALSE,warning=FALSE}
titanic %>% write_csv("titanic_clean.csv")
```
